Uncovering underlying musical content in the temporal domain \\ \small Leveraging deep learning, inductive bias, and aural skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks

### 2023-03-16 introduction/Subsections/Objectives.tex
Updates from Overleaf

>>>>

Let's denote the single embedding for a music piece as $e(M)$.

Given a music piece $M$, we aim to learn a single embedding, $e(M)$, simultaneously capturing the underlying content (melody, harmony, and critical, holistic elements) and the production aspects. The embedding space should be structured so that similar content and production features are clustered together, effectively isolating the two aspects. Formally, our objective is to learn the embedding function $e(M)$ such that:

\begin{itemize}
  \item Similar content features are closer in the embedding space: $\|e(M_1) - e(M_2)\| < \epsilon_c$ for music pieces $M_1$ and $M_2$ with similar content.
  \item Similar production features are closer in the embedding space: $\|e(M_1) - e(M_3)\| < \epsilon_p$ for music pieces $M_1$ and $M_3$ with similar production.
  \item Distinct content and production features are further apart in the embedding space.
\end{itemize}

We describe learning a single embedding $e(M)$ that captures content and production features. The embedding space is structured to cluster similar content and production features together, effectively isolating the two aspects.

### 2023-03-16 methods/subsections/model.tex
Updates from Overleaf

The Triplet Loss function is a function that measures the distance between an anchor, a positive sample, and a negative sample in a high-dimensional space. The formula for this function is given:

$$L = \max( ||f(x_{a}) - f(x_{p})||^{2} - ||f(x_{a}) - f(x_{n})||^{2} + \alpha, 0 )$$

where $f$ is a mapping function that transforms the input samples to a high-dimensional space, $x_{a}$ is the anchor sample, $x_{p}$ is the positive sample, $x_{n}$ is the negative sample, $\alpha$ is a margin hyperparameter that specifies the minimum difference between the distance of the anchor-positive pair and the anchor-negative pair, and $||\cdot||^{2}$ denotes the L2 norm squared.

>>>>

TRIPLET LOSS

 implements the triplet loss function, which is commonly used for learning embeddings in metric learning problems. The triplet loss aims to ensure that the distance between an anchor and a positive sample (both from the same class) is smaller than the distance between the anchor and a negative sample (from a different class) by a margin.

$\mathcal{L} = \frac{1}{N}\sum_{i=1}^{N} \max \left( d\left(a_i, p_i\right) - d\left(a_i, n_i\right) + m, 0 \right)$

In this equation:

$N$ is the number of triplets in the batch
$a_i$ is the $i$-th anchor
$p_i$ is the $i$-th positive sample
$n_i$ is the $i$-th negative sample
$d(\cdot, \cdot)$ is the Euclidean distance function
$m$ is the margin; in this case, it is set to 0.2 by default
The forward method computes the triplet loss as follows:

Calculate the pairwise Euclidean distance between the anchor and positive samples, denoted as $d(a_i, p_i)$.
Calculate the pairwise Euclidean distance between the anchor and negative samples, denoted as $d(a_i, n_i)$.
Compute the loss for each triplet in the batch by subtracting the negative distance from the positive distance and adding the margin: $d(a_i, p_i) - d(a_i, n_i) + m$.
Apply the ReLU function, which is equivalent to taking the maximum value between the computed value and 0: $\max \left( d(a_i, p_i) - d(a_i, n_i) + m, 0 \right)$.
Calculate the mean of the losses across all triplets in the batch.

### 2023-03-16 methods/subsections/test.tex
Updates from Overleaf

We propose the use of a triple siamese network to train a model for music similarity retrieval. Specifically, we aim to minimize the loss function between the anchor, positive, and negative samples by using online triplet mining.

The positive samples are created by transforming and augmenting the original samples while preserving their perceptual underlying musical language. This is achieved by using techniques such as time-stretching, pitch-shifting, and adding noise to the samples. By doing so, we aim to enhance the diversity of the positive samples and improve the model's ability to retrieve similar-sounding samples that share the same underlying musical language.

On the other hand, the negative samples are created by shuffling and scrambling the original samples while maintaining their production texture. This is achieved by using techniques such as random segment selection, time-domain filtering, and frequency-domain masking. The goal is to maintain the production texture of the samples while disrupting their underlying composition.

By using the proposed triple siamese network and online triplet mining, we aim to train a model that can effectively distinguish between similar and dissimilar samples while taking into account both their underlying composition and production texture.

@@

\item Fully connected layers: These layers connect every neuron in one layer to every neuron in another layer, allowing the network to learn non-linear combinations of the features learned in the previous layers.
\vspace*{3mm}

@@

On top of that, one of the main advantages of using CNNs for audio analysis is that they can automatically and adaptively learn temporal hierarchies of features from audio data, which traditional audio processing methods may not be able to do effectively.

>>>>

We propose using a triple siamese network to train a model for music similarity retrieval. Specifically, we aim to minimize the loss function between the anchor, positive, and negative samples by using online triplet mining.

The positive samples are created by transforming and augmenting the original samples while preserving their perceptual underlying musical language. This is achieved using techniques such as time-stretching, pitch-shifting, and adding noise to the samples. By doing so, we aim to enhance the diversity of the positive samples and improve the model's ability to retrieve similar-sounding samples that share the same underlying musical language.

On the other hand, the negative samples are created by shuffling and scrambling the original samples while maintaining their production texture. This is achieved using techniques such as random segment selection, time-domain filtering, and frequency-domain masking. The goal is to maintain the production texture of the samples while disrupting their underlying composition.

Using the proposed triple siamese network and online triplet mining, we aim to train a model that can effectively distinguish between similar and dissimilar samples while considering their underlying composition and production texture.

@@

\item Fully connected layers: These layers connect every neuron in one layer to every neuron in another, allowing the network to learn non-linear combinations of the features learned in the previous layers.
\vspace*{3mm}

@@

On top of that, one of the main advantages of using CNNs for audio analysis is that they can automatically and adaptively learn temporal hierarchies of features from audio data, which traditional audio processing methods may not do effectively.
